### 2022-11-10 paper-thesis/chapters/21-adding-inference.tex
A section about inference

\section{Different sorts of inference}
\label{strength-inference}

The strength of an inference algorithm is how good it is at inferring. Some algorithms are better than others. \cite{typeinference}

\section{Using Statix's first-order unification for inference}

>>>>

\section{Different algorithms for inference}
\label{strength-inference}

There are a lot of different algorithms for inference\cite{typeinference}, some algorithms can solve more inferences than others. However, we would like to avoid implementing an algorithm at all, instead using Statix's built-in first-order unification to do the type inference for us. Implementing an inference algorithm in Statix is theoretically possible but this would be a lot of ugly code, and the goal is to use Statix in a way that is clean and declarative, not to do optimal inference.

The inference that we will build is a very simple unification system, where if at any point during type checking we assert that $\beq{e_1}{e_2}$ and either $e_1$ or $e_2$ is a free variable, we set the free variable to be equal to the other variable. This process is called \emph{first-order unification}.

There are some situations in which this approach fails, but in most real-world scenarios it works perfectly. For example, it can infer both programs in the introduction of this chapter, but it fails to infer the following program:
\begin{lstlisting}
let f = _;
\x : Type.
\g: (_: (f x) -> Bool).
g true
\end{lstlisting}
We know that \verb|f| is a function from \verb|Type -> Type|, but it fails to infer the value of \verb|f|. Because of the way that \verb|g| is used, the type checker asserts that $\beq{f x}{Bool}$. Since x is declared as a function argument and it is completely free, this means that for any \verb|x|, \verb|f x = Bool|. But the rule above is not powerful enough to derive all of this, so it fails.

\section{Using Statix's first-order unification for inference}
\label{statix-inference}
